---
title: "Project-4-Analyst"
author: "Pragya Rawat"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages used 

library(Seurat)
library(dplyr)
library(patchwork)
library(R.filesets)
library(ggplot2)
library(biomaRt)
library(tidyverse)
```
## 4. Load SeuratObject
```{r}
# Load data 

cells <- loadRDS("seurat_3.rds")
```

### 4.1 Identify marker genes for each cluster 
```{r}
# Find markers for every cluster compared to all remaining cells, report only the positive ones

markers <- FindAllMarkers(cells, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
markers <- markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC)
```

```{r}
# Subset for top 10 markers for every cluster with the highest positive log2fc score
top_markers <- markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
```

### 4.2 Label clusters as a cell type based on marker genes
```{r}
# Create subset of clusters acording to marker genes from Supple.Table.S2 

alpha <- subset(markers, gene=="GCG")
beta<- subset(markers, gene=="INS")
delta <- subset(markers, gene=="SST")
gamma <- subset(markers, gene=="PPY")
epsilon <- subset(markers, gene=="GHRL")
ductal<- subset(markers, gene=="KRT19")
acinar<- subset(markers, gene=="CPA1")
stellate <- subset(markers, gene=="PDGFRB")
vasculat <- subset(markers, gene %in% c("VWF","PECAM1","CD34"))
macro <- subset(markers,gene %in% c("CD163", "CD68", "IgG"))
cyto <- subset(markers, gene %in% c("CD3", "CD8"))
mast <-  subset(markers,gene %in% c("TPSAB1", "KIT", "CPA3"))
```
Label the clusters based on marker gene present in each cluster subset:
Beta(INS) - cluster 4
Delta(SST) - cluster 0
Ductal(KRT19) - cluster 9
Gamma(PPY) - cluster 10

```{r}
# Assign Cell Type to Clusters
cell_types <- c("Delta", "1", "2", "3", "Beta", "5", "6", "7", "8", "Ductal", "Gamma", "11", "12", "13")
names(cell_types) <- levels(cells)
cells <- RenameIdents(cells, cell_types)
```
To find unlabeled clusters, use other resources:

1. CELL X GENE: https://cellxgene.cziscience.com/gene-expression
A(TTR) - cluster 2
Mesenchymal(COL1A1, COL3A1, REG3A, REG1B) - cluster 7, cluster 1(low gene expression, high cell count)
Acinar(PRSS2, CTRB2) - cluster 8 
Ductal(TACSTD2, KRT8) - cluster 6 
Endothelial (PLVAP) - cluster 12
Mast(ACP5) - cluster 13 (high gene expression, low cell count)


2. The Human Protein Atlas: https://www.proteinatlas.org/ENSG00000163499-CRYBA2
Alpha (CRYBA2) - cluster 5
Beta (CIQL1, ABCC8) - cluster 11
Ductal (SPTBN2) - cluster 3 
```{r}
# Update cell types 
cell_types <- c("Delta", "Mesenchymal1", "A", "Ductal1", 
                "Beta1", "Alpha", "Ductal2", "Mesenchymal2", 
                "Acinar", "Ductal3", "Gamma", "Beta2", "Endothelial", "Mast")
names(cell_types) <- levels(cells)
cells <- RenameIdents(cells, cell_types)

#check levels
levels(cells)
```

### 4.3 Visualize the clustered cells using a UMAP
```{r}
# Clustered UMAP
png("umap_proj4.png")
DimPlot(cells, reduction = "umap", label = TRUE, pt.size = 0.5) 
dev.off()
```

### 4.4. Visualize the top marker genes in each cluster
```{r}
#Heatmap to visualize top marker genes per cluster 
png("heatmap_proj4.png", height = 1000, width = 1200)
DoHeatmap(cells, features = top_markers$gene, size=3) 
dev.off
```
### 4.5. Find novel marker genes 
```{r}
#Identify novel marker genes
novel_markers <- FindAllMarkers(cells, only.pos = FALSE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r}
# Find significant novel markers with stricter adjusted p value
sig_novel_markers <- novel_markers[novel_markers$p_val_adj<0.005,]
```

### 4.6. Output marker gene information for further downstream analysis 
```{r}
#Save marker genes 
write.csv(markers, "all_markers.csv")
```

```{r}
#Save novel markers
write.csv(sig_novel_markers, "novel_markers.csv")
```

```{r}
#Heatmap of novel genes identified previously
png("heatmap_novel_proj4.png", height = 1000, width = 1200)
top5 <- sig_novel_markers %>% group_by(cluster) %>% top_n(n=5, wt=avg_log2FC)
DoHeatmap(cells, features = top5$gene) 
dev.off()
```

